Article Title

### 2013-12-11 plos.tex
updated introduction

\section*{Abstract}
% Please keep the abstract between 250 and 300 words
Ambitious projects currently under way aim to record the activity of ever larger and denser subsets of neurons \emph{in vivo}.  Correlations measured in such recordings are anticipated to uncover important aspects of the functional organization of neural circuits.  However, estimation and interpretation of large correlation matrices from finite recordings are challenging.  Estimation can be improved by regularization: biasing of the estimate towards a low-dimensional, less variable approximation.  The amount of improvement depends on how parsimoniously the reduced approximation captures the dominant dependencies in the data.  Therefore, the selection of the most efficient estimator is an empirical question that informs about the types of dominant dependencies present in the system.

In this study, we sought the most statistically efficient estimator of neural correlations in recordings from large, dense groups of cortical neurons.  Using fast 3D random-access laser scanning microscopy of calcium signals, we recorded the activity of nearly every neuron in volumes of about 100 $\mu$m in diameter (150--350 cells) in mouse visual cortex.  We hypothesized that, in these dense recordings, the correlation matrix would be most efficiently represented by a sparse network of linear interactions between pairs of observed neurons combined with common latent factors representing unobserved inputs and global network fluctuations.  Indeed, in cross-validation tests, the covariance matrix estimator that imposed this correlation structure outperformed estimators regularized toward other plausible low-dimensional approximations. Furthermore, we found consistent relationships between the organization of sparse interactions inferred by the estimator and the physical distances and preferred orientation differences of pairs of cells.  Consistent with previous synaptic connectivity studies, the density of positive interactions decreased rapidly with distance and with differences in preferred orientation, whereas negative interactions were less selective.  To further corroborate physiological interpretation of the inferred functional structure, future experiments will augment this analysis with measured synaptic connectivity, cell types, and brain states.

@@

By their construction, the correlation between a pair of cells expresses the degree to which the activity of one cell predicts the activity of the other cell. With multiple cells If interactions between variables have a strong linear components, then correlations also statistical dependence.

A correlation matrix provides more information than the equivalent number of pairwise correlations assessed in isolation.  Yet early studies of neural correlations were based on measurements from isolated pairs of neurons; the effects of correlations on neural coding were then extrapolated in simulations and theoretical studies \cite{Shadlen:1998,Zohary:1994}.  Modern multineuronal recordings from increasingly large and dense subsets of neurons in intact circuits allow estimating the entire correlation matrix of large populations of highly interconnected neurons.  Such estimates can be transformed into various representations that express different aspects of the correlation structure and may suggest different mechanistic interpretations.  For example, factor analysis or principal component analysis may reveal common activity fluctuations across the entire population \cite{Yu:2009}.  The correlation matrix also allows computing the \emph{partial correlations} between specific pairs of neurons, \emph{i.e.}\;correlations that remain after accounting for the linear effects of all other recorded cells.  Under several assumptions (predominance of linear effects ( sufficiently complete sampling of interacting neurons, and sufficient sample size), partial correlations may better reflect the direct statistical associations or interactions between pairs of neurons. In multivariate normal distributions, zero partial correlations between pairs of variables indicate conditional independence or lack of interaction between the pair and the network of non-zero interactions constitute Gaussian Graphical Models or Gauss-Markov Random Fields \cite{Koller:2009}. With other distributions, the correspondence between conditional dependence and partial correlations diminishes or breaks down \cite{Loh:2012}. Estimation of networks of partial correlations (sometimes called \emph{association networks}) have been applied to map gene interactions \cite{Schafer:2005,Peng:2009} or to establish functional connectivity between brain regions from fMRI data \cite{Varoquaux:2012,Ryali:2012}.

>>>>

\section*{Abstract}
% Please keep the abstract between 250 and 300 words
Ambitious projects currently under way aim to record the activity of ever larger and denser subsets of neurons \emph{in vivo}.  Correlations measured in such recordings are anticipated to reveal important aspects of the functional organization of neural circuits.  However, estimation and interpretation of large correlation matrices from finite recordings are challenging.  Estimation can be improved by regularization: biasing of the estimate toward a low-dimensional, less variable approximation.  The amount of improvement depends on how parsimoniously the reduced approximation captures the dominant dependencies in the data.  Therefore, the selection of the most efficient estimator is an empirical question that informs about the types of dominant dependencies present in the system.

In this study, we sought the most statistically efficient estimator of neural correlations in recordings from large, dense groups of cortical neurons.  Using fast 3D random-access laser scanning microscopy of calcium signals, we recorded the activity of nearly every neuron in volumes of about 100 $\mu$m in diameter (150--350 cells) in mouse visual cortex.  We hypothesized that, in these dense recordings, the correlation matrix would be most efficiently represented by a sparse network of linear interactions between pairs of observed neurons combined with common latent factors representing unobserved inputs and global network fluctuations.  Indeed, in cross-validation tests, the covariance matrix estimator that imposed this correlation structure outperformed estimators regularized toward other plausible low-dimensional approximations. Furthermore, we found consistent relationships between the organization of sparse interactions inferred by the estimator and the physical distances and preferred orientation differences of pairs of cells.  Consistent with previous synaptic connectivity studies, the density of positive interactions decreased rapidly with distance and with difference in preferred orientation whereas negative interactions were less selective.  To further corroborate physiological interpretation of the inferred functional structure, future experiments will augment this analysis with measured synaptic connectivity, cell types, and brain states.

@@

Yet, a complete correlation matrix provides more information than the equivalent number of isolated pairwise correlations. In early studies, the effects of correlations were extrapolated from isolated pairs to entire populations through simulation and theoretical analysis \cite{Shadlen:1998,Zohary:1994}. In contrast, modern multineuronal recordings from increasingly large and dense populations of highly interconnected neurons allow estimation of various low-dimensional representations of the entire correlation matrix, which can enable inquiry into the nature of underlying physiological mechanisms and computational principles.

In this study, we focused on low-dimensional representations composed of two types of elements: the \emph{low-rank correlation structure} and \emph{partial correlations} between pairs of neurons. The former is most suitable for expressing common circuit fluctuations whereas the latter are most suitable for expressing detailed interactions between specific neurons. Low-rank neural correlation structure extracted by principal component analysis or factor analysis has been analyzed for temporal dynamics in in several studies \cite{Nicolelis:1995,Yu:2006,Yu:2009}.

Partial correlations between pairs of neurons are linear correlations remaining after subtracting the linear effect of all the other neurons. In systems where effects exerted by variables on one another have strong linear components, partial correlations describe conditional dependency: zero  partial correlation between two variables implies conditional independence.  Conditional independence between two variables, if all relevant variables are observed, implies a lack of direct interaction between the underlying processes. Therefore, estimation of networks of partial correlations (sometimes called \emph{association networks}) have been applied to infer interactions between components of complex systems such as gene interaction networks \cite{Schafer:2005,Peng:2009} and functional connectivity between brain regions in fMRI signals \cite{Varoquaux:2012,Ryali:2012}. In multivariate normal distributions\footnote{or any other \emph{elliptical} distribution of the form $g(x-\mu)^\T\Sigma^{-1}(x-\mu))$, where $g$ is an arbitrary function}, are completely defined by linear effects. A multivariate normal distribution defined through a sparse network of non-zero partial correlations is known as a Gaussian Graphical Model or Gauss-Markov Random Field \cite{Koller:2009}. With other distributions where nonlinear effects are more significant, the correspondence between conditional dependence and partial correlations diminishes or breaks down \cite{Loh:2012}.

For example, factor analysis or principal component analysis may reveal common activity fluctuations across the entire population \cite{Yu:2009}.  The correlation matrix also allows computing the \emph{partial correlations} between specific pairs of neurons, \emph{i.e.}\;correlations that remain after accounting for the linear effects of all other recorded cells.  Under several assumptions (predominance of linear effects (sufficiently complete sampling of interacting neurons, and sufficient sample size), partial correlations may better reflect the direct statistical associations or interactions between pairs of neurons. In multivariate normal distributions, zero partial correlations between pairs of variables indicate conditional independence or lack of interaction between the pair and the network of non-zero interactions constitute Gaussian Graphical Models or Gauss-Markov Random Fields \cite{Koller:2009}. With other distributions, the correspondence between conditional dependence and partial correlations diminishes or breaks down \cite{Loh:2012}. Estimation of networks of partial correlations (sometimes called \emph{association networks}) have been applied to map gene interactions \cite{Schafer:2005,Peng:2009} or to establish functional connectivity between brain regions from fMRI data \cite{Varoquaux:2012,Ryali:2012}.
